Tests of Measurement Invariance without Subgroups

### 2012-05-09 m3_2012/invariance.tex
Added draft slides for the m3 conference.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@128 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\title[Measurement Invariance]{Generalized Measurement Invariance
  Tests for Factor Analysis}
\author{Ed Merkle\inst{1} \and Achim Zeileis\inst{2}}
\institute{\inst{1} University of Missouri \and \inst{2} Universit\"{a}t Innsbruck}
\date{Supported by grant SES-1061334 from the U.S.\ National
  Science Foundation}

@@

% FA path diagram, types of MI
\frame{
  \frametitle{Example (Age $\leq$ 16)}
  % path diagram illustrating parameters being the same for everyone
  \begin{center}
      \includegraphics[height=3in]{model_under16.pdf}
  \end{center}}

\frame{
  \frametitle{Example (Age~$>$~16)}
  % path diagram illustrating parameters being the same for everyone
  \begin{center}
      \includegraphics[height=3in]{model_over16.pdf}
  \end{center}}

@@

\frame{
  \frametitle{Lack of Grouping}
  \begin{center}
      \includegraphics[height=3in]{paper-lm-lr-19.pdf}
  \end{center}
}

@@

\frame{
  \frametitle{Simulation}
  \begin{center}
      \includegraphics[height=3in]{paper-mzsim-xyplot.pdf}
  \end{center}
}

@@

\frame{
  \frametitle{Model}
  %\begin{center}
      \includegraphics[height=2.3in]{wicherts_msp.pdf}
  %\end{center}
}

\frame{
  \frametitle{Results for Single Parameters}
  \begin{center}
      \includegraphics[height=3in]{gefp_4_ind.pdf}
  \end{center}}

\frame{
  \frametitle{Aggregated Results}
  \begin{center}
      \includegraphics[height=3in]{gefp_4_agg.pdf}
  \end{center}}

\section{Conclusions}
\frame{
  \frametitle{Conclusions}
  \begin{itemize}
    \item Measurement invariance tests utilizing stochastic processes have 
      important advantages over existing tests:
      \begin{itemize}
        \item Isolating specific parameters that violate
          measurement invariance, allowing the researcher to define
          specific types of measurement invariance ``post hoc''
          instead of ``a priori''.
        \item Isolating
          groups of individuals whose parameter values differ.
        \item Studying the impact of continuous variables on model estimates, without ``ruining'' the rest of the model.
      \end{itemize}
    \item Power is reasonable, with specific tests being better in
      specific circumstances.
  \end{itemize}}

\frame{
  \frametitle{Software}
  \begin{itemize}
    \item To carry out the tests, we utilize
      \begin{itemize}
        \item \texttt{lavaan} for model estimation.
        \item \texttt{estfun()} for score extraction, which is currently a combination of our own code and \texttt{lavaan} code.
        \item \texttt{strucchange} for carrying out the proposed tests with the scores.
          \begin{itemize}
            \item Required input: Fitted model, function for score extraction, and information matrix (optional).
            \item \texttt{gefp()} constructs the process.
            \item \texttt{sctest()} and \texttt{plot()} calculate and visualize test statistics.
          \end{itemize}
      \end{itemize}
  \end{itemize}}

\frame{
  \frametitle{Current Work}
  \begin{itemize}
    \item Continued test implementation via \texttt{strucchange} and \texttt{lavaan} (and possibly \texttt{OpenMx}).\\ \ \\
    \item Detailed examination of test properties.\\ \ \\
    \item Extension to related psychometric issues.\\ \ \\
    \item Working paper: \url{http://econpapers.repec.org/RePEc:inn:wpaper:2011-09}
  \end{itemize}}

>>>>

\title[Measurement Invariance]{Tests of Measurement Invariance without Subgroups}
\author{Ed Merkle\inst{1} \and Achim Zeileis\inst{2}}
\institute{\inst{1} University of Missouri \and \inst{2} Universit\"{a}t Innsbruck}
\date{Supported by grant SES-1061334 from the National
  Science Foundation}

@@

% FA path diagram, types of MI
\frame{
  \frametitle{Example (Age $\leq$ 16)}
  % path diagram illustrating parameters being the same for everyone
  \begin{center}
      \includegraphics[height=3in]{../psychoco2012/model_under16.pdf}
  \end{center}}

\frame{
  \frametitle{Example (Age~$>$~16)}
  % path diagram illustrating parameters being the same for everyone
  \begin{center}
      \includegraphics[height=3in]{../psychoco2012/model_over16.pdf}
  \end{center}}

@@

\frame{
  \frametitle{Lack of Grouping}
  \begin{center}
      \includegraphics[height=3in]{../psychoco2012/paper-lm-lr-19.pdf}
  \end{center}
}

@@

\frame{
  \frametitle{Simulation}
  \begin{center}
      \includegraphics[height=3in]{../psychoco2012/paper-mzsim-xyplot.pdf}
  \end{center}
}

@@

\frame{
  \frametitle{Model}
  %\begin{center}
      \includegraphics[height=2.3in]{../psychoco2012/wicherts_msp.pdf}
  %\end{center}
}

\frame{
  \frametitle{Results for Single Parameters}
  \begin{center}
      \includegraphics[height=3in]{../psychoco2012/gefp_4_ind.pdf}
  \end{center}}

\frame{
  \frametitle{Aggregated Results}
  \begin{center}
      \includegraphics[height=3in]{../psychoco2012/gefp_4_agg.pdf}
  \end{center}}

\section{Conclusions}
\frame{
  \frametitle{Conclusions}
  \begin{itemize}
    \item Measurement invariance tests utilizing stochastic processes have 
      important advantages over existing tests:
      \begin{itemize}
        \item Isolating specific parameters that violate
          measurement invariance, allowing the researcher to interpret the nature of measurement invariance violations.
        \item Isolating
          groups of individuals whose parameter values differ.
        \item Studying the impact of continuous variables on model estimates, without ``ruining'' the rest of the model.
      \end{itemize}
    \item Power is reasonable, with specific tests being better in
      specific circumstances.
  \end{itemize}}

\frame{
  \frametitle{Software}
  \begin{itemize}
    \item We can carry out the tests for general SEMs in R, using
      \begin{itemize}
        \item \texttt{lavaan} for model estimation.
        \item \texttt{estfun()} for score extraction, which is currently a combination of our own code and \texttt{lavaan} code.
        \item \texttt{strucchange} for carrying out the proposed tests with the scores.
          % \begin{itemize}
          %   \item Required input: Fitted model, function for score extraction, and information matrix (optional).
          %   \item \texttt{gefp()} constructs the process.
          %   \item \texttt{sctest()} and \texttt{plot()} calculate and visualize test statistics.
          % \end{itemize}
      \end{itemize}
  \end{itemize}}

\frame{
  \frametitle{Current Work}
  \begin{itemize}
    \item Continued test implementation via \texttt{strucchange} and \texttt{lavaan} (and possibly \texttt{OpenMx}).\\ \ \\
    \item Detailed examination of test properties.\\ \ \\
    \item Extension to related psychometric issues.\\ \ \\
    \item Paper: Merkle, E.\ C.\ \& Zeileis, A.\ (accepted).  Tests of measurement invariance without subgroups:\ {A} generalization of classical methods.  {\em Psychometrika}.
  \end{itemize}}
